What's hot and what's not:\\Windowing developer topic analysis

### 2009-04-11 lda-paper.tex
editted the intro

  As development on a software project progresses, developers will
  naturally shift their focus between different topics and tasks many
  times.  Managers and newcomer developers often seek ways of
  understanding what tasks have recently been worked on and how much
  effort has gone into each; for example, a manager might wonder what
  unexpected tasks occupied their team's attention during a period
  when they were supposed to have been implementing a set of new
  features.  Tools such as Latent Dirichlet Allocation (LDA) and
  Latent Semantic Indexing (LSI) can be used to analyze check-in
  comments over the entire lifetime of a project, and previous work on
  developer topic analysis has leveraged these tools to associate
  commit log comments with independent topics extracted from these
  commit log comments.  In this work, we use LDA and LSI to analyze
  periods, such as months, within a project's lifetime to create a
  time-windowed model of changing development topics.  We propose a
  visualization of this model that allows us to explore the evolving
  stream of topics of development occurring over time.  We demonstrate
  that windowed topic analysis offers advantages over topic analysis
  applied to a projects lifetime because many topics are quite local.

@@

Imagine that a development team have agreed upon a set of features to
implement for their next iteration, but these features were not
finished by the end of the iteration. The developers insist they were
focussed on the features, yet their manager and is not sure what
happened. In a scenario like this, topic analysis can help tease out
the topics of development that the developers were focussed on. Topic
analysis could highlight the orthogonal issues that developers were
spending time on.  Topic analysis extracts groups of words which from
documents (commit log messages) which can be used to relate
documents. These extracted topics often correlate with actual
development topics that the developers discussed. By applying topic
analysis locally to an iteration, both the developers and managers
might be able to see what were the topics of development.

Commits in SCSs can have
multiple purposes. Thus we hope that our research will help uncover
these multiple purposes that manifest themselves as topics interleaved
within the commits.  Others have demonstrated that 
topic analysis techniques, such as Latent Semantic Indexing (LSI) and
Latent Dirichlet Allocation (LDA), can aid in separating topics from
these mixed comments. These topics often describe distinct ideas that
are interleaved throughout the development. A topic represents a
groups of commit comments that are related to each other by their
content.  In this case a topic is a set of tokens.

@@

We propose to extend existing topic analysis techniques by
applying them to windows of documents over time. We hope by looking
at topic clusters during periods or windows over time we can identify
ongoing areas of focus on a select group of topics as well reflect more intermittent topics. Figure \ref{fig:lda} illustrates
the kind of visualization we want to automatically generate. We want
to see topics that are unique to a certain time period, displayed
across the time axis. Those topics that recur, or occur over a
larger period are plotted continuously. In our example we have
titled each topic with a single word.

In this paper we explore how topics shift over time in the source
control repository of a project, using several open source database
systems as examples. We analyze commit comments per revision in each
time window and we extract the topics in that time window. We expect
that topics will change over time and the similarities
and differences in these topics will indicate developer
focus and changes in developer focus as it changes over time.

Our contributions include:
\begin{itemize}
\item Demonstrating the value of windowed topic analysis using trends
\item Multiple visualizations topics and trends over time
\item An exploratory case study using these techniques on several open source database systems.
\end{itemize}

>>>>

  As development on a software project progresses, developers will
  naturally shift their focus between different topics and tasks many
  times.  Managers and newcomer developers often seek ways of
  understanding what tasks have recently been worked on and how much
  effort has gone into each; for example, a manager might wonder what
  unexpected tasks occupied their team's attention during a period
  when they were supposed to have been implementing a set of new
  features.  Tools such as Latent Dirichlet Allocation (LDA) and
  Latent Semantic Indexing (LSI) can be used to analyze check-in
  comments over the entire lifetime of a project. Previous work on
  developer topic analysis has leveraged these tools to associate
  commit log comments with independent topics extracted from these
  commit log comments.  In this work, we use LDA and LSI to analyze
  periods, such as months, within a project's lifetime to create a
  time-windowed model of changing development topics.  We propose a
  visualization of this model that allows us to explore the evolving
  stream of topics of development occurring over time.  We demonstrate
  that windowed topic analysis offers advantages over topic analysis
  applied to a projects lifetime because many topics are quite local.

@@

Imagine that a development team had agreed upon a set of features to
implement for a development iteration, but these features were not
finished by the end of the iteration. The developers insist they
focussed on those features, yet their manager is not sure what
happened. In a scenario like this, topic analysis can help tease out
the topics of development that the developers were focussed on. Topic
analysis could highlight the orthogonal issues that developers were
spending time on.  Topic analysis extracts groups of words from
documents (commit log messages) that are word distributions, topics,
which characterize clusters of documents. These extracted topics often
correlate with actual development topics that the developers
discussed. By applying topic analysis locally to an iteration, both
the developers and managers might be able to see what were the topics
of development during that iteration. %XXX RM during that iteration

Commits in SCSs can have
multiple purposes. Thus we hope that our research will help uncover
these multiple purposes that manifest themselves as topics interleaved
within the commits.  Others have demonstrated that 
topic analysis techniques, such as Latent Semantic Indexing (LSI) and
Latent Dirichlet Allocation (LDA), can aid in separating topics from
these mixed comments. These topics often describe distinct ideas that
are interleaved throughout the development. A topic represents a
groups of commit log comments that are related to each other by their
content.  In this case a topic is a set of tokens.

@@

We propose to extend existing topic analysis techniques by
applying them to windows of documents over time. We hope that by
looking %XX X ADDED
at topic clusters during periods or windows over time we can identify
ongoing areas of focus on a select group of topics as well reflect more intermittent topics. Figure \ref{fig:lda} illustrates
the kind of visualization we want to automatically generate. We want
to see topics that are unique to a certain time period, displayed
across the time axis. Those topics that recur, or occur over a
larger period are plotted continuously. In our example we have
titled each topic with a single word.

In this paper we explore how topics shift over time in the source
control system (SCS) of a project, using several open source database
systems as examples. We analyze commit log comments in each
time window and we extract the topics of that time window. We expect
that topics will change over time and the similarities and differences
in these topics will indicate developer focus and changes in developer
focus as it changes over time.

Our contributions include:
\begin{itemize}
\item Demonstrating the value of windowed topic analysis using trends.
\item Multiple visualizations of topics and trends over time.
\item An exploratory case study using these techniques on several open source database systems.
\end{itemize}
